~\\ \textbf{\laType\ \laAssignment: \laTitle \tab }

\normalsize 

    \subsection*{Information}
    
        In-class labs are meant to introduce you to a new topic and provide
        some practice with that new topic. 
    
        \paragraph{Topics:} \laTopics
        
        \paragraph{Solo work:}
        Labs should be worked on by each
        individual student, though asking others for help is permitted.
        \underline{Do not} copy code from other sources, and do not give your
        code to other students.
        \textbf{ \color{red} Students who commit or aid in plagiarism will receive a 0\% on the assignment and be reported. }

        \paragraph{Building and running:} If you are using Visual Studio,
            make sure to run \underline{with} debugging.
            \textbf{ \color{red} Do not run without debugging! }

            ~\\
            Using the debugger will help you find errors. \\
            To prevent a program exit, use this before \texttt{return 0;}

            \begin{mdframed} \centering
                \texttt{ cin.ignore(); \tab cin.get(); }
            \end{mdframed}

        \paragraph{Turn in:}
            Once you're ready to turn in your code, prepare the files by
            doing the following: \textbf{(1)} Make a copy of your project folder and name it \texttt{LASTNAME-FIRSTNAME-LABNAME}.
            \textbf{(2)} Make sure that all source files (.cpp, .hpp, and/or .h files) and the \texttt{Makefile} files
            are all present.
            \textbf{(3)} Remove all Visual Studio files - I only want the source files and Makefiles.
            \textbf{(4)} Zip your project folder as \texttt{LASTNAME-FIRSTNAME-LABNAME.zip}
            
            ~\\ \textbf{ \color{red} Never turn in Visual Studio files! }

        \paragraph{Starter files:} \laStarterFiles

        \paragraph{Grading:}
            Grading is based on completion, if the program functions as intended,
            and absense of errors. \textbf{ \color{red} Programs that don't build will receive 0\%. }
            Besides build errors, runtime errors, logic errors, memory leaks, and ugly code will reduce your score.


        \tableofcontents

    \newpage
