\documentclass[a4paper,12pt]{book}
\usepackage[utf8]{inputenc}
\title{}
\author{Rachel Morris}
\date{\today}

\usepackage{rachwidgets}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{dirtree}
\usepackage{boxedminipage}

\newcommand{\laClass}{CS 210\ }
\newcommand{\laSemester}{Fall 2017\ }

\newcounter{question}

\pagestyle{fancy}
\fancyhf{}
\lhead{\laClass}
\chead{\laSemester}
\rhead{Exam 1 \testSet}
\rfoot{\thepage\ of \pageref{LastPage}}
\lfoot{\scriptsize By Rachel Morris, last updated \today}

\renewcommand{\headrulewidth}{2pt}
\renewcommand{\footrulewidth}{1pt}

\begin{document}

    \chapter*{Exam 1 Review}
    
    The exam has 7 questions over the following content. It will be given
    in the testing center at JCCC over several days.

    \begin{enumerate}
        \item How is an STL vector implemented behind the scenes? \\
            See: http://www.cplusplus.com/reference/vector/vector/

        \item How is an STL list implemented behind the scenes? \\
            See: http://www.cplusplus.com/reference/list/list/
        
        \item Make sure you can write the class declarations
            for Singly Linked List and Doubly Linked List Nodes.
            
        \item What is a Circularly Linked List?
        
        \item What does a pointer point to when not initialized? What should
            it be initialized to in order to avoid undefined behavior in
            your program?
            
        \item How do you delete a dynamic variable? How do you delete a dynamic array?
        
        \item Make sure to be able to find $O(n)$ values for basic functions.
        
        \item Make sure you know how to implement functionality from a DoublyLinkedList, including:
            
            \begin{itemize}
                \item   PushFront
                \item   PushBack
                \item   PopFront
                \item   PopBack
                \item   GetAtIndex
            \end{itemize}
            
            You will have to implement 3 of them on paper for the exam.
                   
    \end{enumerate}










\end{document}
