\documentclass[a4paper,12pt]{book}
\usepackage[utf8]{inputenc}
\title{}
\author{Rachel Morris}
\date{\today}

\usepackage{rachwidgets}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{dirtree}
\usepackage{boxedminipage}

\newcommand{\laClass}{CS 210\ }
\newcommand{\laSemester}{Fall 2017\ }

\newcounter{question}

\pagestyle{fancy}
\fancyhf{}
\lhead{\laClass}
\chead{\laSemester}
\rhead{Exam 2 Review}
\rfoot{\thepage\ of \pageref{LastPage}}
\lfoot{\scriptsize By Rachel Morris, last updated \today}

\renewcommand{\headrulewidth}{2pt}
\renewcommand{\footrulewidth}{1pt}

\begin{document}

    \chapter*{Exam 2 Review}
    
    The exam has 12 questions over the following content. It will be given
    in the testing center at JCCC over several days.

    ~\\
    You can bring a calculator in order to do modulus operations - non-graphing only!

    ~\\
    This exam does not have code-writing problems like the first exam.

    \paragraph{Polymorphism}

    \begin{enumerate}
        \item   Use the lecture slides from http://edu.moosader.com/course/cs200/ for reference.
        \item   What is Polymorphism? What is its purpose?
        \item   How do you use Polymorphism in the code?
        \item   How are pointers used in Polymorphism?
        \item   How is Polymorphism related to a family of classes?
        \item   How much functions be declared in the family to ensure that
                    the correct function is called when using Polymorphism?
    \end{enumerate}

    \paragraph{Stacks and Queues}

    Questions are similar to the Stack and Queue lab...

    \begin{enumerate}
        \item   What are the common add, access, and remove functions
                    for Stacks and for Queues?
        \item   What is ``FIFO'' and ``LIFO'', and which structures does each belong to?
        \item   Given a list of items, what is the result after Pop() for a Stack and for a Queue?
        \item   Draw a diagram of a Stack and a Queue after a series of Push/Pop calls.
        \item   Given a list of items, how many Pop() commands are needed to access an item, for both a Stack and a Queue.
    \end{enumerate}

    \paragraph{Dictionaries}

    \begin{enumerate}
        \item   Be familiar with how to compute Modulus.
        \item   For some key and some TABLE\_SIZE, be able to compute an index
        \item   Be able to identify when a generated index creates a collision
        \item   Be able to step through basic Linear Probing functionality (step forward by 1)
        \item   Be able to step through basic Quadratic Probing functionality (step forward by $i^{2}$)
        \item   Given a Hash and a Hash2 function, be able to calculate an index using Double Hashing.
    \end{enumerate}

    \paragraph{Algorithm Efficiency}

    \begin{enumerate}
        \item   Use http://bigocheatsheet.com/ for studying from.   \\
                (A Dictionary is also known as a Hash Table)
        \item   Know the average efficiency for...:
        \begin{itemize}
            \item   Adding an item to the end of an array
            \item   Adding an item to the end of a linked list
            \item   Adding an item to a dictionary
            \item   Accessing an item at some arbitrary position in an array
            \item   Accessing an item at some arbitrary position in a linked list
            \item   Searching through an unsorted array for a value
            \item   Searching through a linked list for a value
        \end{itemize}
    \end{enumerate}










\end{document}
